The Strong Gravitational Lens Finding Challenge

### 2019-02-17 paper.tex
Edit

In deriving a good figure of merit for evaluating lens finding algorithms one needs to take into account the particular nature of this problem.
The traditional method for evaluating a classification algorithm is with the receiver operating characteristic curve, or  ROC  curve.  This is a plot of the true positive rate (TPR) versus the false positive rate (FPR).  In this case these are defined as
\begin{align}
{\rm TPR} &= \frac{\textrm{ number of true lenses classified as lenses}}{\textrm{ total number of true lenses}} \\
{\rm FPR} &= \frac{\textrm{number of non-lenses classified as lenses}}{\textrm{ total number of non-lenses}}
\end{align}
The classifier generally gives a probability of a candidate being a lens, $p$, in which case a threshold is set and everything with $p$ greater is classified as a lens and everything smaller is classified as not a lens.  The TPR and FPR are then plotted as a curve parametrised by this threshold.  At $p=1$ all of the cases are classified as non-lenses and so TPR=FPR=1 and at $p=0$ all of the cases are classified as lenses so TPR=FPR=0.  These points are always added to the ROC curve.  If the classifier made random guesses then the ratio of lenses to non-lenses would be the same as the ratio of the number of cases classified as lens to the number of cases classified as non-lenses and so TPR=FPR.  The better a classifier is the smaller the FPR and the larger the TPR so the further away from this diagonal line it will be.  When a classifier provides only a binary classification or a discrete ranking, the ROC connects the endpoints to the discrete points found by using each rank as a threshold.

@@

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/einstein_space.pdf}
 \caption{The AUROC, TPR$_0$, TPR$_{10}$ and the fraction of lenses in the test sample after discarding the lenses with Einstein radii larger than the number indicated on the x-axis. }
 \label{fig:einstein_space}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/einstein_ground.pdf}
 \caption{Same as figure~\ref{fig:einstein_space}, but for space-based entries.}
 \label{fig:einstein_ground}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/flux_space.pdf}
 \caption{Same as figure~\ref{fig:einstein_space}, but here the x-axis is the flux within the pixels that are above 1 $\sigma$ in the lensed source only image.}
 \label{fig:flux_space}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/flux_ground.pdf}
 \caption{Same as figure~\ref{fig:flux_space}, but for space-based entries.}
 \label{fig:flux_ground}
\end{figure*}

\begin{figure*}
\includegraphics[width=2\columnwidth]{figures/npixel_space.pdf}
 \caption{Same as figure~\ref{fig:einstein_space}, but here the x-axis is the number of pixels that are above 1 $\sigma$ in the lensed source only image.  This is an indication of the lensed arcs' size.}
 \label{fig:npixel_space}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/npixel_ground.pdf}
 \caption{Same as figure~\ref{fig:npixel_space}, but for space-based entries.}
 \label{fig:npixel_ground}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/flux_contrast_space.pdf}
  \caption{Same as figure~\ref{fig:einstein_space}, but here the x-axis is the ratio of the flux coming from the lensed source to the total flux in the image in the index band.}
 \label{fig:flux_contrast_space}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/flux_contrast_ground.pdf}
 \caption{Same as figure~\ref{fig:flux_contrast_space}, but for space-based entries.}
 \label{fig:flux_contrast_ground}
\end{figure*}

>>>>

In deriving a good figure of merit for evaluating lens finding algorithms one needs to take into account the particular nature of this problem.
The traditional method for evaluating a classification algorithm is with the receiver operating characteristic curve, or  ROC  curve.  This is a plot of the true positive rate (TPR) versus the false positive rate (FPR).  In this case these are defined as
\begin{align}
{\rm TPR} &= \frac{\textrm{ number of true lenses classified as lenses}}{\textrm{ total number of true lenses}} \\
{\rm FPR} &= \frac{\textrm{number of non-lenses classified as lenses}}{\textrm{ total number of non-lenses}}
\end{align}
The classifier generally gives a probability of a candidate being a lens, $p$, in which case a threshold is set and everything with $p$ greater is classified as a lens and everything smaller is classified as not a lens.  The TPR and FPR are then plotted as a curve parametrised by this threshold.  At $p=1$ all of the cases are classified as non-lenses and so TPR=FPR=0 and at $p=0$ all of the cases are classified as lenses so TPR=FPR=1.  These points are always added to the ROC curve.  If the classifier made random guesses then the ratio of lenses to non-lenses would be the same as the ratio of the number of cases classified as lens to the number of cases classified as non-lenses and so TPR=FPR.  The better a classifier is the smaller the FPR and the larger the TPR so the further away from this diagonal line it will be.  When a classifier provides only a binary classification or a discrete ranking, the ROC connects the endpoints to the discrete points found by using each rank as a threshold.

@@

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/einstein_space.pdf}
 \caption{The AUROC, TPR$_0$, TPR$_{10}$ and the fraction of lenses in the test sample after discarding the lenses with Einstein radii larger than the number indicated on the x-axis. The vertical dotted lines indicate where no more than 100 lenses in the test sample had larger Einstein radii.  Beyond this point one should be suspicious of small number statistics.  }
 \label{fig:einstein_space}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/einstein_ground.pdf}
 \caption{Same as figure~\ref{fig:einstein_space}, but for ground-based entries.  The AUROC, TPR$_0$, TPR$_{10}$ and the fraction of lenses in the test sample after discarding the lenses with Einstein radii larger than the number indicated on the x-axis. The vertical dotted lines indicate where no more than 100 lenses in the test sample had larger Einstein radii.  }
 \label{fig:einstein_ground}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/flux_space.pdf}
 \caption{ The AUROC, TPR$_0$, TPR$_{10}$ and the fraction of lenses in the space-based test sample after discarding the lenses with fluxes within the pixels that are above 1 $\sigma$ in the lensed source as indicated on the x-axis. The vertical dotted lines indicate where no more than 100 lenses in the test sample had larger Einstein radii. }
 \label{fig:flux_space}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/flux_ground.pdf}
 \caption{Same as figure~\ref{fig:flux_space}, but for ground-based entries.}
 \label{fig:flux_ground}
\end{figure*}

\begin{figure*}
\includegraphics[width=2\columnwidth]{figures/npixel_space.pdf}
 \caption{The AUROC, TPR$_0$, TPR$_{10}$ and the fraction of lenses as a function of the lens image size for the space-based test set.  The x-axis is the number of pixels that are above 1 $\sigma$ in the lensed source only image.  This is an indication of the lensed arcs' size.  The vertical dotted lines indicate where no more than 100 lenses in the test sample had larger Einstein radii. 
 }
 \label{fig:npixel_space}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/npixel_ground.pdf}
 \caption{Same as figure~\ref{fig:npixel_space}, but for ground-based entries.}
 \label{fig:npixel_ground}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/flux_contrast_space.pdf}
  \caption{Same as figure~\ref{fig:flux_contrast_space}, but here the x-axis is the ratio of the flux coming from the lensed source to the total flux in the image in the index band.  This is for the space-based test set.}
 \label{fig:flux_contrast_space}
\end{figure*}

\begin{figure*}
 \includegraphics[width=2\columnwidth]{figures/flux_contrast_ground.pdf}
 \caption{Same as figure~\ref{fig:flux_contrast_space}, but for ground-based entries.}
 \label{fig:flux_contrast_ground}
\end{figure*}
